% Informace o práci/projektu / Information about the thesis
%---------------------------------------------------------------------------
\projectinfo{
%
%Prace / Thesis
    project={BP},            %typ práce BP/SP/DP/DR  / thesis type (SP = term project)
    year={2023},             % rok odevzdání / year of submission
    date=\today,             % datum odevzdání / submission date
%
%Nazev prace / thesis title
    title.cs={Využití zpětnovazebného učení pro automatickou alokaci akciového portfolia},  %  thesis title in czech language (according to assignment)
    title.en={Reinforcement Learning for Automated Stock Portfolio Allocation}, %  thesis title in english
    title.length={14.5cm}, %  setting the length of a block with a thesis title for adjusting a line break (can be defined here or below)
    sectitle.length={14.5cm}, %  setting the length of a block with a second thesis title for adjusting a line break (can be defined here or below)
    dectitle.length={14.5cm}, %  setting the length of a block with a thesis title above declaration for adjusting a line break (can be defined here or below)
%
%Autor / Author
    author.name={Zdeněk},   % jméno autora / author name
    author.surname={Lapeš},   % příjmení autora / author surname
%
%Ustav / Department
    department={UITS}, % UIFS/UITS/UPGM / fill in appropriate abbreviation of the department according to assignment: UPSY/UIFS/UITS/UPGM
%
% Školitel / supervisor
    supervisor.name={Milan},   % jméno školitele / supervisor name
    supervisor.surname={Češka},   % příjmení školitele / supervisor surname
    supervisor.title.p={doc. RNDr.},   %titul před jménem (nepovinné) / title before the name (optional)
    supervisor.title.a={Ph.D.},    %titul za jménem (nepovinné) / title after the name (optional)
%
% Klíčová slova / keywords
    keywords.cs={
        umělá inteligence, AI, posilované učení, alokace akciového portfolia,
        moderni teorie portfolia, Q-learning, neuronové sítě, akciový trh
    },
    keywords.en={
        artificial intelligence, AI, reinforcement learning, stock portfolio allocation,
        modern portfolio theory, Q-learning, neural networks, stock market
    },
%
% Abstrakt / Abstract
    abstract.cs={
        Tato prace se zameruje na navrh a implementaci modelu schopnych
        delat rozhodnuti ve velmi rychle se menicim financnim prostredi. Vychazi z teorie
        moderniho portfolia a posilovaneho uceni. V prvnim kroku je vytvoreno
        simulacni prostredi, ktere umoznuje testovat ruzne strategie. V druhem kroku
        je trenovan agent, ktery se snazi dosahnout maximalniho navratu.
        V tretim kroku je agent testovan na realnych datech.
    }, % abstrakt v českém či slovenském jazyce / abstract in czech or slovak language
    abstract.en={
        This thesis focuses on the design and implementation of a model capable of
        making decisions in a very fast-changing financial environment. It is based on the
        modern portfolio therory and reinforcement learning.
        In the first step, a simulation environment is created
        that allows testing various strategies. In the second step
        the agent is trained to try to achieve the maximum return.
        In the third step, the agent is tested on real data.
    }, % abstrakt v anglickém jazyce / abstract in english
%  Declaration (for thesis in english should be in english; for project practice can be commented out)
    declaration={
        I hereby declare that this Bachelor's thesis was prepared as an original work by the author under
        the supervision of Mr. Milan Češka, Ph.D. % TODO: make this dynamic
        I have listed all the literary sources, publications and other sources, which were used during the preparation of this thesis.
    },
%
%  Acknowledgement (optional, ideally in the language of the thesis)
    acknowledgment={
        I would like to thank my supervisor, Mr. Milan Češka, for his guidance and support during the preparation of this thesis.
        I would also like to thank my family and friends for their support.
    },
%
%  Extended abstract (approximately 3 standard pages) - can be defined here or below
%    extendedabstract={ % path must be like that because this is inserted directly into xlapes02.tex
%        \input{/Users/zlapik/my-drive-zlapik/0-todo/ai-investing/docs/content/extended_abstract.tex}
%    },
%
    faculty={FIT}, % FIT/FEKT/FSI/FA/FCH/FP/FAST/FAVU/USI/DEF
}
